Multilevel spin models in cold atomic systems

### 2021-04-15 sun_phases/spin_model_paper/main.tex
'finalized' main body of SU(n) spin model paper (prior to review/feedback)

% more shorthands
\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}

\newcommand{\xx}{\x\x}
\newcommand{\xxi}{\x\x_\i}
\newcommand{\X}{\text{X}}
\newcommand{\XX}{\X\X}
\newcommand{\XXI}{\X\X_\i}
\newcommand{\su}{\mathfrak{su}}
\newcommand{\MF}{\text{MF}}
\renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
\DeclareMathOperator{\diag}{diag}

%%% for in-line enumerated lists
\usepackage[inline]{enumitem}
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering
\setlist{nolistsep} % more compact spacing between environments

@@

If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $c_{\pi-p,\mu}^\dag c_{\pi-q,\mu} c_{p\nu}^\dag c_{q\nu}$, which conserve both momentum and energy.
We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally-symmetric spin states with one atom per lattice site.}.
The operator content of terms with $k=\ell$ and $p=q$ takes the form $c_{\ell\mu}^\dag c_{\ell\mu} c_{q\nu}^\dag c_{q\nu}$, so these terms contribute an overall energy $\frac12NU$ that can be neglected in the absence of coherent atom number fluctuations.
Defining the spin operators $s_{\mu\nu q}\equiv c_{q\mu}^\dag c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians become
\begin{align}
  H_{\t{kin}} &= -J\sum_{q,\mu} \cos\p{q} s_{\mu\mu q},
  \label{H_kin_start} \\
  H_{\t{int}} &= -\f{U}{2N} \sum_{p,q,\mu,\nu} s_{\mu\nu p} s_{\nu\mu q}.
  \label{eq:H_int_start}
\end{align}
Throughout this work, we will assume that atomic modes are singly-occupied, e.g.~due to the initialization of a spin-polarized state with one atom per lattice site, in which multiple occupation of an atomic mode is forbidden by fermionic statistics (Pauli exclusion).
In this case, we can neglect the fermionic origins of $H_{\t{int}}$ and instead treat our system as a bona fide spin model.
Note that the ``kinetic'' terms of this spin model ($H_{\t{kin}}$) are proportional to the identity operator, and can be neglected for the time being.
The fermionic origin of these terms will return to play an important role when we consider spin-orbit coupling in Section \ref{sec:SOC}.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{spin_model.pdf}
\caption{
(a) Ultracold atoms on a lattice of $L$ sites can tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U_0$.
(b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
(d) Interactions thereby open an energy gap $U=U_0\times N/L$ between the manifold of permutationally-symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
}
\label{fig:spin_model}
\end{figure}

We now discuss the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
The operator $\v s_p\c\v s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
The term $-\v s_p\c\v s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally-symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
See Figure \ref{fig:spin_model} for a summary of this section thus far.

@@

External fields or additional interactions that respect permutational symmetry can induce nontrivial dynamics within the PS manifold.
Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small (compared to the energy gap $U$) these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}).
Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.

@@

In principle, making the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $H_{\t{int}} \to H_{\t{int}}^{(\phi)}$, where
\begin{align}
  H_{\t{int}}^{(\phi)}
  \equiv -\f{U}{2N} \sum_{p,q,\mu,\nu}
  c_{p-\mu\phi,\mu}^\dag c_{p-\nu\phi,\nu}
  c_{q-\nu\phi,\nu}^\dag c_{q-\mu\phi,\mu}.
\end{align}
However, $H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $H_{\t{kin}}^{(\phi)}$.
Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ ($J$ large compared to $U$) while keeping $J\phi$ constant.
Altogether, the interacting spin Hamiltonian in the gauged frame becomes
\begin{align}
  H_{\t{spin}} = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q},
  \label{eq:H_spin}
\end{align}
consisting of a spin-locking $\v S\c\v S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.

@@

Eqs.~\eqref{eq:drive_all} and \eqref{eq:drive_single} are the main technical results of this section.
There are three important observations to make about these results.
First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}.
Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-s_\alpha^2$ and $\pm s_\alpha+s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\c\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(s_\x,s_\y,s_\z)$.
The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{S_{\mu\mu}}$ (where $S_{\mu\nu}=\sum_js_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v S}=\sum_{\mu\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes $\Omega_m$.

Finally, we comment on the preparation of initial states.
Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v_+$, i.e.~$\ket{\v v_+}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauged frame).
In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
For example, when $n$ is even we can prepare an $N$-fold product of the state
\begin{align}
  \frac{\ket{s} + \ket{-s}}{\sqrt{2}}
  \simeq e^{-\i\frac{\pi}{2}\p{s_\y + s_\y^2}} \ket{s},
\end{align}
where $\simeq$ denotes equality up to a global phase.
This state has a vanishing mean spin vector, $\bk{s_\x}=\bk{s_\y}=\bk{s_\z}=0$, but variances $\bk{s_\x^2}=\bk{s_\y^2}=s/2$ and $\bk{s_\z^2}=s^2$.

@@

We now return to the spin Hamiltonian $H_{\t{spin}}$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $H_{\t{spin}}$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
\begin{align}
  H_\MF = \sum_q\sp{-U\bk{\bar{\v s}}\c\v s_q
    + J\phi \sum_q \sin\p{q} s_{\z,q}}.
  \label{eq:H_MF}
\end{align}
where $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix.
We assume that all momenta $q\in\ZZ_N\times 2\pi/N$ are occupied.
Fixing the atom number $N$, the spin Hamiltonian has one free parameter: the reduced field $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
One should therefore expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $J\phi/U\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.

@@

Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\X\equiv\ket\x^{\otimes N}$, where
\begin{align}
  \ket\x \equiv e^{-\i\frac{\pi}{2}s_\y} \ket{s}
  = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
  \label{eq:state_x}
\end{align}
Here ${ m \choose k }$ is a binomial coefficient.
As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0^+$.
The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
By minimizing the reduced field $J\phi/U$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).

\begin{figure}
\centering
\includegraphics{crit_fields_X.pdf}
\caption{
The critical value of $\p{J\phi/U}_{\t{crit}}$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
A single-parameter fit to $\p{J\phi/U}_{\t{crit}}=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
}
\label{fig:crit_fields_X}
\end{figure}

As shown in insets of Figure \ref{fig:mean_mag_int_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
\begin{align}
  \sigma_\MF \to \f{\sigma_\MF}{\gamma\p{n/2}},
  &&
  \bbk{\ss}_\MF \to \f{\bbk{\ss}_\MF-\gamma\p{n}}{1-\gamma\p{n}},
  \label{eq:rescale}
\end{align}
where
\begin{align}
  \gamma\p{k} \equiv \f{\Gamma\p{k-\frac12}}{\sqrt\pi\,\Gamma\p{k}}
  \stackrel{k\ge2}{\approx} \f1{\sqrt{\pi(k-1)}}.
  \label{eq:gamma}
\end{align}
The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0^+$.

@@

We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
\begin{align}
  \ket\xx \equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}},
  &&
  \ket\xxi \equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}},
\end{align}
and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
\begin{align}
  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}
  = \f1{2^s} \sum_\mu \p{-1}^{s+\mu}
  { 2s \choose s+\mu }^{1/2} \ket{\mu}.
\end{align}
Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} respectively show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_{\t{spin}}$ and $\ket\xxi=e^{-\i\frac{\pi}{2}s_\z}\ket\x\equiv\ket\y$.
The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.

>>>>

% more shorthands
\newcommand{\up}{\uparrow}
\newcommand{\dn}{\downarrow}
\newcommand{\x}{\text{x}}
\newcommand{\y}{\text{y}}
\newcommand{\z}{\text{z}}
\newcommand{\g}{\text{g}}
\newcommand{\e}{\text{e}}
\renewcommand{\SS}{\mathbb{S}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\A}{\mathcal{A}}
\newcommand{\C}{\mathcal{C}}
\newcommand{\D}{\mathcal{D}}
\newcommand{\E}{\mathcal{E}}
\newcommand{\I}{\mathcal{I}}
\renewcommand{\O}{\mathcal{O}}
\renewcommand{\P}{\mathcal{P}}

\newcommand{\xx}{\x\x}
\newcommand{\xxi}{\x\x_\i}
\newcommand{\X}{\text{X}}
\newcommand{\XX}{\X\X}
\newcommand{\XXI}{\X\X_\i}
\newcommand{\spin}{\text{spin}}
\newcommand{\eff}{\text{eff}}
\newcommand{\crit}{\text{crit}}
\newcommand{\MF}{\text{MF}}
\newcommand{\su}{\mathfrak{su}}
\renewcommand{\ss}{\bar{\v s}\c\bar{\v s}}
\DeclareMathOperator{\diag}{diag}

\newcommand{\col}{\underline}
\newcommand{\mean}{\overline}

%%% for in-line enumerated lists
\usepackage[inline]{enumitem}
\setlist[enumerate,1]{label={(\roman*)}} % default in-line numbering

@@

If the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, then the mode-changing collisions in $H_{\t{int}}$ become off-resonant, motivating the frozen-mode approximation $\set{k,p}=\set{\ell,q}$ (which we assume for the remainder of this work)\footnote{Note that the frozen-mode approximation neglects correlated momentum-hopping terms of the form $c_{\pi-p,\mu}^\dag c_{\pi-q,\mu} c_{p\nu}^\dag c_{q\nu}$, which conserve both momentum and energy.
We defer a careful treatment of these terms to future work, noting only that they vanish on the manifold of permutationally symmetric spin states with one atom per lattice site.}.
The operator content of terms with $k=\ell$ and $p=q$ takes the form $c_{\ell\mu}^\dag c_{\ell\mu} c_{q\nu}^\dag c_{q\nu}$, so these terms contribute an overall energy $\frac12NU$ that can be neglected in the absence of coherent atom number fluctuations.
Defining the spin operators $s_{\mu\nu q}\equiv c_{q\mu}^\dag c_{q\nu}$, the remaining terms of the kinetic and interaction Hamiltonians become
\begin{align}
  H_{\t{kin}} &= -J\sum_{q,\mu} \cos\p{q} s_{\mu\mu q},
  \label{H_kin_start} \\
  H_{\t{int}} &= -\f{U}{2N} \sum_{p,q,\mu,\nu} s_{\mu\nu p} s_{\nu\mu q}.
  \label{eq:H_int_start}
\end{align}
Throughout this work, we will assume that atomic modes are singly-occupied, e.g.~due to the initialization of a spin-polarized state with one atom per lattice site, in which multiple occupation of an atomic mode is forbidden by fermionic statistics (Pauli exclusion).
In this case, we can neglect the fermionic origins of $H_{\t{int}}$ and instead treat our system as a bona fide spin model.
Note that the ``kinetic'' terms of this spin model ($H_{\t{kin}}$) are proportional to the identity operator, and can be neglected for the time being.
The fermionic origin of these terms will return to play an important role when we consider spin-orbit coupling in Section \ref{sec:SOC}.

@@

\begin{figure}
\centering
\includegraphics[width=\linewidth]{spin_model.pdf}
\caption{
(a) Ultracold atoms on a lattice of $L$ sites can tunnel between neighboring lattice sites at a rate $J$, and locally repel each other with interaction energy $U_0$.
(b,c) When the interaction energy $U_0$ is smaller than the single-particle bandwidth $\sim J$, the frozen-mode approximation enables writing the interaction Hamiltonian as a spin model consisting of exchange terms $\v s_p\c\v s_q$, which swap the states two spins pinned to momentum modes $p,q$.
(d) Interactions thereby open an energy gap $U=U_0\times N/L$ between the manifold of permutationally symmetric states of $N$ spins, and the orthogonal complement of states that break spin-permutation symmetry.
}
\label{fig:spin_model}
\end{figure}

We now discuss the spin Hamiltonian $H_{\t{int}}$ in Eq.~\eqref{eq:H_int}.
The operator $\v s_p\c\v s_q$ simply swaps the nuclear spin states of two atoms pinned to modes $p,q$.
The term $-\v s_p\c\v s_q$ thereby assigns a definite energy of $-1$ ($+1$) to a pair of spins that are symmetric (anti-symmetric) under exchange.
In this sense, $\v s_p\c\v s_q$ is analogous to the enforcement of SU(2) spin alignment by ferromagnetic interactions, which similarly assigns (distinct) definite energies to anti-symmetric spin-0 singlet $\ket{\up\dn}-\ket{\dn\up}$ and the symmetric spin-1 triplets $\set{\ket{\up\up},\ket{\dn\dn},\ket{\up\dn}+\ket{\dn\up}}$.
By summing over all pair-wise exchange terms $\v s_p\c\v s_q$, the interaction Hamiltonian $H_{\t{int}}$ energetically enforces a permutational symmetry among all spins, opening an energy gap $U$ between the manifold of all permutationally symmetric (PS) states and the orthogonal complement of excited (e.g.~spin-wave) states that break permutational symmetry.
See Figure \ref{fig:spin_model} for a summary of this section thus far.

@@

External fields or additional interactions that respect permutational symmetry can induce nontrivial dynamics within the PS manifold.
Even if additional Hamiltonian terms break permutational symmetry, as long as the coupling to states outside of the PS manifold is small (compared to the energy gap $U$) these terms can be restricted to the PS manifold perturbatively (see Appendix \ref{sec:pert_theory}) \cite{bravyi2011schrieffer}.
Simulating dynamics within the PS manifold requires calculating matrix elements $\bk{\ell|\O|m}$ of spin operators $\O$ with respect to PS states $\ket\ell,\ket{m}$; we discuss this calculation in Appendix \ref{sec:PS_ops}.

@@

In principle, making the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ also transforms the interaction Hamiltonian as $H_{\t{int}} \to H_{\t{int}}^{(\phi)}$, where
\begin{align}
  H_{\t{int}}^{(\phi)}
  \equiv -\f{U}{2N} \sum_{p,q,\mu,\nu}
  c_{p-\mu\phi,\mu}^\dag c_{p-\nu\phi,\nu}
  c_{q-\nu\phi,\nu}^\dag c_{q-\mu\phi,\mu}.
\end{align}
However, $H_{\t{int}}^{(\phi)}$ cannot be exactly reduced to a spin Hamiltonian that addresses the same spin degrees of freedom as $H_{\t{kin}}^{(\phi)}$.
Nonetheless, for sufficiently weak SOC (small $s\phi$) the effect of $H_{\t{int}}^{(\phi)}$ may be approximated by $H_{\t{int}}^{(0)}=H_{\t{int}}$, which has previously been benchmarked for SU(2)-symmetric interactions \cite{he2019engineering, smale2019observation}.
To ensure that $H_{\t{kin}}^{(\phi)}$ does not become trivial as $\phi\to0$, we can simultaneously take $J\to\infty$ ($J$ large compared to $U$) while keeping $J\phi$ constant.
Altogether, the interacting spin Hamiltonian in the gauged frame becomes
\begin{align}
  H_\spin = -\f{U}{2N} \v S\c\v S + J\phi \sum_q \sin\p{q} s_{\z,q},
  \label{eq:H_spin}
\end{align}
consisting of a spin-locking $\v S\c\v S$ term that energetically favors permutational symmetry, and an inhomogeneous magnetic field that causes inter-spin dephasing.

@@

Eqs.~\eqref{eq:drive_all} and \eqref{eq:drive_single} are the main technical results of this section.
There are three important observations to make about these results.
First, the fact that $H_{\t{3LD}}$ acts identically on all spins means we can freely replace the site index $j$ with a momentum index $q$ (as can be verified by substituting $c_{j\mu}=\frac1{\sqrt{L}}\sum_k e^{-\i q\c j} c_{q\mu}$), which is important to ensure that this drive addresses the same spin degrees of freedom as the spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}.
Second, each of $\tilde\Omega_0,\tilde\Omega_+,\tilde\Omega_-$ can be tuned independently by changing the amplitudes of the driving lasers; some particular Hamiltonians for specific values of these amplitudes are shown in Table \ref{tab:drives}.
Third, due to the appearance of mutually commuting pairs of Hamiltonians in Table \ref{tab:drives}, specifically $-s_\alpha^2$ and $\pm s_\alpha+s_\alpha^2$ for $\alpha\in\set{\z,\x}$, the three-laser drive admits pulse sequences that exactly implement arbitrary SU(2) (spatial) rotations of the form $e^{-\i\chi\vec n\c\vec s}$, where $\chi$ is a rotation angle, $\vec n$ is a rotation axis, and $\vec s\equiv(s_\x,s_\y,s_\z)$.
The capability to perform arbitrary spatial rotations, together with the capability to measure the number of atoms with spin projection $\mu$ onto a fixed quantization axis, $\bk{S_{\mu\mu}}$ (where $S_{\mu\nu}=\sum_js_{\mu\nu j}$), implies the capability to reconstruct all components of the mean collective spin matrix $\bk{\v S}=\sum_{\mu\nu}\bk{S_{\mu\nu}}\op{\mu}{\nu}$ via spin qudit tomography \cite{newton1968measurability, perlin2020qudit}.
Moreover, we expect that advanced quantum optimal control techniques (similar to those of Refs.~\cite{anderson2015accurate, lucarelli2018quantum}) can be used to implement arbitrary SU($n$) rotations by designing suitable time-dependent drive amplitudes $\Omega_m$.

Finally, we comment on the preparation of initial states.
Initial states are nominally prepared in the ``lab frame'', and must be transformed according to the gauge transformation $c_{q\mu}^\dag\to c_{q-\mu\phi,\mu}^\dag$ prior to evolution under the ``gauged frame'' spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin} or the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
We assume the capability to prepare an initial state in which all spins are maximally polarized along $\v v_+$, i.e.~$\ket{\v v_+}^{\otimes N}=\ket{s}^{\otimes N}$, which is unaffected by the gauge transformation (up to a global phase).
The three-laser then allows us to rotate this state into one that is polarized along any spatial axis (in the gauged frame).
In addition, when $n>2$ the three-laser drive allows us to prepare product states with nontrivial intra-spin correlations.
For example, when $n$ is even we can prepare an $N$-fold product of the state
\begin{align}
  \frac{\ket{s} + \ket{-s}}{\sqrt{2}}
  \simeq e^{-\i\frac{\pi}{2}\p{s_\y + s_\y^2}} \ket{s},
\end{align}
where $\simeq$ denotes equality up to a global phase.
This state has a vanishing mean spin vector, $\bk{s_\x}=\bk{s_\y}=\bk{s_\z}=0$, but variances $\bk{s_\x^2}=\bk{s_\y^2}=s/2$ and $\bk{s_\z^2}=s^2$.

@@

We now return to the spin Hamiltonian $H_\spin$ in Eq.~\eqref{eq:H_spin}, which is valid in the weak-SOC limit $s\phi\ll1$, and henceforth work exclusively in the ``gauged frame'' of $H_\spin$ and the three-laser drive $H_{\t{3LD}}$ in Eq.~\eqref{eq:drive_all}.
At the mean-field (MF) level, up to constant energy shifts the undriven spin Hamiltonian becomes
\begin{align}
  H_\MF = \sum_q\sp{-U\bk{\bar{\v s}}\c\v s_q
    + J\phi \sum_q \sin\p{q} s_{\z,q}}.
  \label{eq:H_MF}
\end{align}
where $\bar{\v s}\equiv\frac1N\sum_q\v s_q$ is the average spin matrix.
We assume that all momenta $q\in\ZZ_N\times 2\pi/N$ are occupied.
Fixing the atom number $N$, the spin Hamiltonian has one free parameter: the reduced field $J\phi/U$ that determines the relative strength of the single-particle and interaction terms.
One should therefore expect distinct dynamical behaviors when $J\phi/U\ll1$, in which case strong spin-locking interactions should give rise to a long-range ordered phase, as opposed to $J\phi/U\gg1$, in which case long-range order should be destroyed by the strong inhomogeneous magnetic field.

@@

Figure \ref{fig:mean_mag_int_X} shows the time-averages of the magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ as computed by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\X\equiv\ket\x^{\otimes N}$, where
\begin{align}
  \ket\x \equiv e^{-\i\frac{\pi}{2}s_\y} \ket{s}
  = \f1{2^s} \sum_\mu { 2s \choose s+\mu }^{1/2} \ket{\mu}.
  \label{eq:state_x}
\end{align}
Here ${ m \choose k }$ is a binomial coefficient.
As expected, the spin model exhibits a mean-field dynamical phase transition between an ordered phase at small $J\phi/U$ and a disordered phase at large $J\phi/U$.
The ordered phase has a non-zero magnetization $\sigma_\MF$ and an interaction energy $\bbk{\ss}_\MF$ that asymptotically approach their maximal values as $J\phi/U\to0$.
The disordered phase has no (time-averaged) magnetization, $\sigma_\MF=0$, but the interaction energy $\bbk{\ss}_\MF$ nonetheless indicates persistent nontrivial inter-spin correlations when $n>2$.
These nontrivial correlations vanish as $J\phi/U\to\infty$, in which case $\bbk{\ss}_\MF$ approaches the minimal value allowed by conservation laws (clarified below).
By minimizing the reduced field $J\phi/U$ for which $\sigma_\MF=0$, we numerically find that the transition between ordered and disordered phases occurs at a critical field $\p{J\phi/U}_\crit=\p{n/2}^{-\alpha}$ with $\alpha\approx1/3$ (see Figure \ref{fig:crit_fields_X}).
When $n=2$, this transition is consistent with the predictions of a Lax vector analysis \cite{yuzbashyan2005nonequilibrium, yuzbashyan2006dynamical, yuzbashyan2006relaxation, yuzbashyan2015quantum, smale2019observation} that exploits integrability of $H_\spin$ to determine long-time behavior.
However, a straightforward application of the same analysis fails to predict the mean-field transition for $n>2$, so additional theoretical tools are necessary to understand this transition.
We elaborate on this point in Appendix \ref{sec:lax}.

\begin{figure}
\centering
\includegraphics{crit_fields_X.pdf}
\caption{
The critical value of $\p{J\phi/U}_\crit$ as determined by mean-field simulations of $N=100$ spins initially in the $x$-polarized state $\ket\X$.
A single-parameter fit to $\p{J\phi/U}_\crit=\p{n/2}^{-\alpha}$ finds $\alpha=0.333(5)$, and $\alpha=1/3$ is consistent with all mean-field results to within an uncertainty determined by the resolution of $J\phi/U$ in mean-field simulations.
}
\label{fig:crit_fields_X}
\end{figure}

As shown in insets of Figure \ref{fig:mean_mag_int_X}, mean-field results for different spin dimensions $n$ collapse onto each other when normalizing the field $J\phi/U$ to its critical value, $\p{J\phi/U}\to\p{J\phi/U}\times\p{n/2}^{1/3}$, and rescaling
\begin{align}
  \sigma_\MF \to \f{\sigma_\MF}{\gamma\p{n/2}},
  &&
  \bbk{\ss}_\MF \to \f{\bbk{\ss}_\MF-\gamma\p{n}}{1-\gamma\p{n}},
  \label{eq:rescale}
\end{align}
where
\begin{align}
  \gamma\p{k} \equiv \f{\Gamma\p{k-\frac12}}{\sqrt\pi\,\Gamma\p{k}}
  \stackrel{k\ge2}{\approx} \f1{\sqrt{\pi(k-1)}}.
  \label{eq:gamma}
\end{align}
The rescaling of magnetization and interaction energy can be understood by considering their limiting behavior as $J\phi/U\to\infty$ or $J\phi/U\to0$.

@@

We now consider the same setup as above, but with the initial ``multi-cat'' states $\ket\XX\equiv\ket\xx^{\otimes N}$ and $\ket\XXI\equiv\ket\xxi^{\otimes N}$, where
\begin{align}
  \ket\xx \equiv \f{\ket{\x} + \ket{-\x}}{\sqrt{2}},
  &&
  \ket\xxi \equiv \f{\ket{\x} + \p{-1}^s \ket{-\x}}{\sqrt{2}},
\end{align}
and $\ket{-\x}$ is a state polarized along $-x$, defined similarly to $\ket\x$ in Eq.~\eqref{eq:state_x}:
\begin{align}
  \ket{-\x} \equiv e^{-\i\frac{\pi}{2}s_\y}\ket{-s}
  = \f1{2^s} \sum_\mu \p{-1}^{s+\mu}
  { 2s \choose s+\mu }^{1/2} \ket{\mu}.
\end{align}
Similarly to Figure \ref{fig:mean_mag_int_X}, Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} show the time-averaged magnetization $\sigma_\MF$ and interaction energy $\bbk{\ss}_\MF$ throughout mean-field dynamics of the initial states $\ket\XX$ and $\ket\XXI$.
These figures exclude the trivial case of spin dimension $n=2$, for which $\ket\xx=\ket{-s}$ is an eigenstate of $H_\spin$ and $\ket\xxi=e^{-\i\frac{\pi}{2}s_\z}\ket\x\equiv\ket\y$ is spin-polarized along $y$.
The first and perhaps most interesting observation to make about Figures \ref{fig:mean_mag_int_XX} and \ref{fig:mean_mag_int_XXI} is that they are different, signifying the importance of intra-spin coherences for the dynamical behavior of multilevel spin models.
